Estimating dark matter halo concentration using the integrated mass profile

### 2015-11-09 paper/paper.tex
abstract + fluido

Galaxy properties are coupled to the structural of its dark
matter halo host. 
One of the key quantities is the halo density profile which, assuming
a  Navarro-Frenk-White parameterization, can be described by the
concentration.
This has motivated the development of different algorithms to estimate
the concentraion in N-body halos.
Here we present a new algorithm to estimate the concentration
based on the radially integrated mass profile
The main advantage of this approach is that it uses the full
particle information without any binning, which makes it reliable in
cases when low numerical resolution becomes a limitation for other
methods.   
We show the good performance of this method by estimating halo
concentration both on mock and N-body halos. 
We compare these results against two other methods:
maximum radial velocity measurements and radial particle binning to
fit the density profile. 
Tests on  the mock halos show that the accuracy of our method to recover input
concentrations varies with halo resolution. For halos
sampled with $20$ particles our method recovers the input
concentration with $10\%$ accuracy. 
For the maximum radial velocity method it is on the order of $20\%$
while density profile methods perform worst at the low resolution.
For halos with $2\times 10^4$ particles our method achieves an
accuracy of $0.01\%$ while the velocity and density profile methods
achieve $0.1\%$ and $1\%$, respectively. 
We measure the mass-concentration relationship on N-body
data taking care of using halos sampled with at least $200$
particles. We find that at low masses $10^{12}\hMsun <
M<10^{13}\hMsun$ our method yields median concentration values lower
by a $15\%-20\%$ compared to the velocity and density methods 
At higher masses $3\times 10^{13}\hMsun < M < 2\times
10^{14}\hMsun$ the three methods give similar results.
\end{abstract}

@@

Let us review first the basic properties of the NFW density profile.
This shall help us to define our notation and review the important
assumptions of the two different methods commonly used to measure the
concentration to finally present the new method.

@@

\begin{equation}
\rho(r) = \frac{\rho_c\delta_c}{r/r_s(1+r/r_s)^2},
\label{eq:definition}
\end{equation}
%
where $\rho_c\equiv 3H^2/8\pi G$ is the Universe critical density, $H$
is the Hubble constant, $G$ and the universal gravitational constant,
$\delta_c$ is the halo dimensionless characteristic density and $r_s$
is the scale radius. 
This radius marks the point where the logarithmic slope of the density
profile is equal to -2, marking the transition between the power law
scaling $\rho\propto r^{-1}$ for $r<r_s$ and $\rho\propto r^{-3}$ for
$r>r_s$.

@@

In this paper we use $\Delta_h=740$ roughly corresponding to 200 times
the critical density at redshift z=0.

@@

>>>>

Galaxy properties are coupled to the structural of its dark
matter halo host. 
One of the key quantities is the halo density profile which, assuming
a  Navarro-Frenk-White parameterization, can be described by the
concentration.
This has motivated the development of different algorithms to estimate
the concentraion in N-body halos.
Here we present a new algorithm to estimate the concentration
based on the radially integrated mass profile
The algorithm is easy to implement. It main advantages is that uses the full
particle information without any binning, making it realiable in
cases when low numerical resolution becomes a limitation for other
methods. 
We show the good performance of this method by estimating halo
concentration both on mock and N-body halos. 
We compare these results against two other methods:
maximum radial velocity measurements and radial particle binning to
fit the density profile. 
Tests on  the mock halos show that the accuracy of our method to recover input
concentrations varies with halo resolution. For halos
sampled with $20$ particles our method recovers the input
concentration with $10\%$ accuracy. 
For the maximum radial velocity method it is on the order of $20\%$
while density profile methods perform worst at the low resolution.
For halos with $2\times 10^4$ particles our method achieves an
accuracy of $0.01\%$ while the velocity and density profile methods
achieve $0.1\%$ and $1\%$, respectively. 
We measure the mass-concentration relationship on N-body
data taking care of using halos sampled with at least $200$
particles. We find that at low masses $10^{12}\hMsun <
M<10^{13}\hMsun$ our method yields median concentration values lower
by a $15\%-20\%$ compared to the velocity and density methods 
At higher masses $3\times 10^{13}\hMsun < M < 2\times
10^{14}\hMsun$ the three methods give similar results.
\end{abstract}

@@

Let us review first the basic properties of the NFW density profile.
This shall help us to define our notation.

@@

\begin{equation}
\rho(r) = \frac{\rho_c\delta_c}{r/r_s(1+r/r_s)^2},
\label{eq:definition}
\end{equation}
%
where $\rho_c\equiv 3H^2/8\pi G$ is the Universe critical density, $H$
is the Hubble constant, $G$ and the universal gravitational constant,
$\delta_c$ is the halo dimensionless characteristic density and $r_s$
is the scale radius. 
This radius marks the point where the logarithmic slope of the density
profile is equal to -2, the transition between the power law
scaling $\rho\propto r^{-1}$ for $r<r_s$ and $\rho\propto r^{-3}$ for
$r>r_s$.

@@

In this paper we use $\Delta_h=740$, a number roughly corresponding to
200 times the critical density at redshift z=0.

@@

We now review the key assumptions of the different methods used
to estimate the concentration.
